\documentclass[a4paper,UKenglish]{darts-v2019}
 
\usepackage{microtype}

\nolinenumbers

\bibliographystyle{plainurl}

\newenvironment{scope}{\section{Scope}}{}
\newenvironment{content}{\section{Content}}{}
\newenvironment{getting}{\section{Getting the artifact} The artifact 
endorsed by the Artifact Evaluation Committee is available free of 
charge on the Dagstuhl Research Online Publication Server (DROPS).}{}
\newenvironment{platforms}{\section{Tested platforms}}{}
\newcommand{\license}[1]{{\section{License}#1}}
\newcommand{\mdsum}[1]{{\section{MD5 sum of the artifact}#1}}
\newcommand{\artifactsize}[1]{{\section{Size of the artifact}#1}}


% Author macros::begin %%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%

\title{Don't Panic! Better, Fewer, Syntax Errors for LR Parsers: Artefact}

\author{Lukas Diekmann}{Software Development Team, King's College London, United Kingdom \and \url{https://lukasdiekmann.com/}}{lukas.diekmann@gmail.com}{}{}
\author{Laurence Tratt}{Software Development Team, King's College London, United Kingdom \and \url{https://tratt.net/laurie/}}{laurie@tratt.net}{https://orcid.org/0000-0002-5258-3805}{}
\authorrunning{L.\,Diekmann and L.\,Tratt}
\Copyright{Lukas Diekmann and Laurence Tratt}
\keywords{Parsing, error recovery, programming languages}
\ccsdesc{Theory of computation~Parsing}
\ccsdesc{Software and its engineering~Compilers}

\RelatedArticle{<placeholder>}

\Volume{6}
\Issue{2}
\Article{<placeholder>}
\RelatedConference{34th European Conference on Object-Oriented Programming (ECOOP 2020), July 13--17, 2020, Berlin, Germany}

\begin{document}

\maketitle

\begin{abstract}
This is the artefact accompanying the paper `Don't Panic! Better, Fewer, Syntax
Errors for LR Parsers` by Diekmann and Tratt. It focusses on the experiment
from that paper, which compares a number of different error recovery algorithms
on a large corpus of data, including all the software necessary to reproduce
the experiment from the paper.
\end{abstract}

\begin{scope}
The artefact focusses on the experiment in Section 6 from the paper. It
includes the software necessary to build and run the experiment. The experiment
makes use of Java source code with syntax errors from
\href{https://bluej.org/blackbox/}{Blackbox}, which we are not allowed to
distribute. Instead we include anonymised identifiers which allow those who
\href{https://bluej.org/blackbox/#Access}{register with Blackbox} to precisely
recreate the corpus we used. If you do not wish to register with Blackbox, you
can use your own corpus of Java source code instead.
\end{scope}

\begin{content}
The artefact package includes:
\begin{itemize}
\item The software necessary to build and run the experiment.
\item The specific version of the \emph{grmtools} parsing software (created for
    the paper) used in the experiment.
\end{itemize}
\end{content}

\begin{getting}
In addition, the artefact is also available at:
\url{https://archive.org/download/error_recovery_experiment/0.4/}
\end{getting}

\begin{platforms}
The experiment is likely to work on any platform on which Rust 1.42 (or later)
and Python 2.7 run. The artefact \texttt{README.md} includes further details
about additional software dependencies.
\end{platforms}

\license{The artefact is available under both the Apache License (Version 2.0)
and the MIT License.}

\mdsum{070d328911ec6e9b400f4d2db82b61f3}

\artifactsize{5.7 MiB}

\subparagraph*{Acknowledgements.}

We are grateful to the Blackbox developers for allowing us access to their data
and tooling. This research was funded by the EPSRC Lecture (EP/L02344X/1)
Fellowship.

\end{document}
